---
title: "Penguin Bill Anatomy"
author: "Nidhi Danda"
date: "2026-09-23"
format: html
---

## Introduction

Data: [Palmer Penguins](https://allisonhorst.github.io/palmerpenguins/), In this post, I will look at bill length and bill depth, and how they differ between the 3 species. I will also examine if you can determine a species from its bill shape alone. 

```{r}
#| message: false
#| warning: false
library(palmerpenguins)
library(tidyverse)
```

## Examining the data

```{r}
penguins |>
    drop_na() |>
    glimpse()
```

## Summary statistics

```{r}
#| label: tbl-bill-summary
#| tbl-cap: "Mean bill length and depth by species."

penguins |>
    drop_na() |>
    group_by(species) |>
    summarise(
        mean_bill_length = mean(bill_length_mm), 
        mean_bill_depth = mean(bill_depth_mm),
        n = n()
    )
```

## Visualising bill shape by species
```{r}
#| label: fig-penguin-bill
#| fig-cap: "Bill length vs bill depth, coloured by species"

penguins |>
    drop_na() |>
    ggplot(aes(x = bill_length_mm, y = bill_depth_mm, colour = species)) + 
    geom_point(alpha = 0.7) + 
    labs(x = "Bill length(mm)", y = "Bill depth (mm)") + 
    theme_minimal()
```

## Findings

Adelie has shorter bill lengths but larger bill depths, whilst Gentoo has larger bill lengths and shorter bill depths. However, Chinstrap has relatively large bill depth and bill length. 